---
title: "Math 308: Managing API Data"
author: "Laesch, Subbiah, Louchiem, Tantum"
date: "Spring 2023"
output:
  pdf_document: default
  html_document: default
---

```{r, setup, include=FALSE}
require(mosaic)
library(formatR)
library(tidycensus)
library(tidyverse)
library(rvest)
library(gridExtra)
library(sf)
knitr::opts_chunk$set(
  tidy.opts=list(width.cutoff=80), 
  tidy=TRUE)   # slightly smaller font for code
```

## Data Overview:

1. Median Income by tract (Census, pulled) 
2. Poverty rate of high school students by tract (Census, pulled)
3. Median income by educational attainment level (Census, unpulled)  
4. Open choice assignment by attendance areas (Seattle Schools Report, unpulled)  
5. Median income by mobility (Census, pulled)  
6. Achievement results by school/by demographic (USDoE, pulled)


Generate API Key for Census Data Scraping:

```{r, message=FALSE, warning = FALSE}
census_api_key("3488cfc51ef04c9886a285b2a7c770d4dcff78f8")

# To install API key for use in future sessions, run this function with `install = TRUE`
```

2010 Data: Median Income and Percent Under Poverty:

```{r, message=FALSE, warning=FALSE}
# acs1vars <- load_variables(year = 2010, dataset = "acs1") # Costly run time
medianIncomeVar = c("B19013_001")

# Median Household Income: 2006-2010 5-year ACS
acs1 = get_acs(geography = "tract", variables = medianIncomeVar, state = "53", county = "033", year = 2010, geometry = TRUE)
attach(acs1)

# Read in list of Seattle block groups from SeattleGeoData
seattletracts <- read.csv("~/R/MATH 308 Data/Census_Tracts_2010.csv")  %>%
  pull("GEOID10") # Extracts a single column

# Subset data to Seattle Tracts
data10 <- acs1 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition

# Make Map (Change Estimate Scale)
data10 %>%
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c(option = "magma") +
  ggtitle("King County By Median Household Income") +
  labs(fill = "Income Amount\n(2010 Census)")

seattleIndexGeoId <- read.csv("~/R/MATH 308 Data/Racial_and_Social_Equity_Composite_Index_Current.csv") %>%
  pull("GEOID")

dataTrial <- data10 %>%
  filter(GEOID %in% seattleIndexGeoId) %>%
  pull(GEOID)

seattleIndex <- read.csv("~/R/MATH 308 Data/Racial_and_Social_Equity_Composite_Index_Current.csv")
newIndex <- seattleIndex %>%
  filter(GEOID %in% dataTrial) %>%
  pull(SOCIOECON_DISADV_PERCENTILE)

newData <- data10 %>%
  filter(GEOID %in% seattleIndexGeoId)

newData$Poverty_Rate <- newIndex


newData %>%
  ggplot(aes(fill = Poverty_Rate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c(option = "magma") +
  ggtitle("Percent Living under 200% Poverty in King County") +
  labs(fill = "Poverty Rate\n(2010 Census)")
```

Faceted Mapping: 2010

```{r, message=FALSE}
# Median Income: Same house 1 year ago
medianIncomeByMobility1 = c("B07011_002")
# Median Income: Moved within the same county
medianIncomeByMobility2 = c("B07011_003")
# Median Income: Moved from a different county within the same state
medianIncomeByMobility3 = c("B07011_004")
# Median Income: Moved from a different state
medianIncomeByMobility4 = c("B07011_005")

# Median Household Income: 2006-2010 5-year ACS
acsMob1 = get_acs(geography = "tract", variables = medianIncomeByMobility1, state = "53", county = "033", year = 2010, geometry = TRUE)
attach(acsMob1)

acsMob2 = get_acs(geography = "tract", variables = medianIncomeByMobility2, state = "53", county = "033", year = 2010, geometry = TRUE)
attach(acsMob2)

acsMob3 = get_acs(geography = "tract", variables = medianIncomeByMobility3, state = "53", county = "033", year = 2010, geometry = TRUE)
attach(acsMob3)

acsMob4 = get_acs(geography = "tract", variables = medianIncomeByMobility4, state = "53", county = "033", year = 2010, geometry = TRUE)
attach(acsMob4)

# Subset data to Seattle Tracts
dataMob1 <- acsMob1 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition
dataMob2 <- acsMob2 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition
dataMob3 <- acsMob3 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition
dataMob4 <- acsMob4 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition

# Make Map (Change Estimate Scale)
p1 <- dataMob1 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Same house as previous year\n") +
  labs(fill = "Income Amount\n(2010 Census)")

p2 <- dataMob2 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Moved within the same county\n") +
  labs(fill = "Income Amount\n(2010 Census)")

p3 <- dataMob3 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Moved within the same state\n") +
  labs(fill = "Income Amount\n(2010 Census)")

p4 <- dataMob4 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Moved from a different state\n") +
  labs(fill = "Income Amount\n(2010 Census)")

grid.arrange(p1, p2, p3, p4)
```  
Faceted Mapping: -- Scale is weird

```{r, message=FALSE, error=FALSE, warning=FALSE}
# Median Income: Same house 1 year ago
medianIncomeByMobility1 = c("B07011_002")
# Median Income: Moved within the same county
medianIncomeByMobility2 = c("B07011_003")
# Median Income: Moved from a different county within the same state
medianIncomeByMobility3 = c("B07011_004")
# Median Income: Moved from a different state
medianIncomeByMobility4 = c("B07011_005")

# Median Household Income: 2006-2010 5-year ACS
acsMob1_15 = get_acs(geography = "tract", variables = medianIncomeByMobility1, state = "53", county = "033", year = 2015, geometry = TRUE)
attach(acsMob1_15)

acsMob2_15 = get_acs(geography = "tract", variables = medianIncomeByMobility2, state = "53", county = "033", year = 2015, geometry = TRUE)
attach(acsMob2_15)

acsMob3_15 = get_acs(geography = "tract", variables = medianIncomeByMobility3, state = "53", county = "033", year = 2015, geometry = TRUE)
attach(acsMob3_15)

acsMob4_15 = get_acs(geography = "tract", variables = medianIncomeByMobility4, state = "53", county = "033", year = 2015, geometry = TRUE)
attach(acsMob4_15)

# Subset data to Seattle Tracts
dataMob1_15 <- acsMob1_15 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition
dataMob2_15 <- acsMob2_15 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition
dataMob3_15 <- acsMob3_15 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition
dataMob4_15 <- acsMob4_15 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition

# Make Map (Change Estimate Scale)
d1 <- dataMob1_15 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Same house as previous year\n") +
  labs(fill = "Income Amount\n(2015 Census)")

d2 <- dataMob2_15 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Moved within the same county\n") +
  labs(fill = "Income Amount\n(2015 Census)")

d3 <- dataMob3_15 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Moved within the same state\n") +
  labs(fill = "Income Amount\n(2015 Census)")

d4 <- dataMob4_15 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Moved from a different state\n") +
  labs(fill = "Income Amount\n(2015 Census)")

grid.arrange(d1, d2, d3, d4)
```  

Comparing Changing Median Household Income for Long Term Residents:

```{r, message=FALSE, error=FALSE, warning=FALSE}
acsMob1_17 = get_acs(geography = "tract", variables = medianIncomeByMobility1, state = "53", county = "033", year = 2017, geometry = TRUE)
attach(acsMob1_17)
dataMob1_17 <- acsMob1_17 %>%
  filter(GEOID %in% seattletracts)

b1 <- dataMob1_17 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Same house as previous year\n") +
  labs(fill = "Income Amount\n(2017 Census)")

acsMob1_19 = get_acs(geography = "tract", variables = medianIncomeByMobility1, state = "53", county = "033", year = 2019, geometry = TRUE)
attach(acsMob1_19)
dataMob1_19 <- acsMob1_19 %>%
  filter(GEOID %in% seattletracts)

c1 <- dataMob1_19 %>% 
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Household Income by Mobility", subtitle = "Same house as previous year\n") +
  labs(fill = "Income Amount\n(2019 Census)")

grid.arrange(p1, d1, b1, c1, widths = c(1.5,2))
```

Poverty Rates of High School Students:  

```{r, message=FALSE, error=FALSE, warning=FALSE}
# High School Enrollment: Above Poverty Line
schoolPov1 = c("B14006_018")
# High School Enrollment: Below Poverty Line  
schoolPov2 = c("B14006_008")

# High School Enrollment: Above Poverty Line: 2006-2010 5-year ACS
acsSchoolPov1 = get_acs(geography = "tract", variables = schoolPov1, state = "53", county = "033", year = 2010, geometry = TRUE)
attach(acsSchoolPov1)

# High School Enrollment: Below Poverty Line
acsSchoolPov2 = get_acs(geography = "tract", variables = schoolPov2, state = "53", county = "033", year = 2010, geometry = TRUE)
attach(acsSchoolPov1)

# Subset data to Seattle Tracts
dataSPov1 <- acsSchoolPov1 %>%
  filter(GEOID %in% seattletracts)
dataSPov2 <- acsSchoolPov2 %>%
  filter(GEOID %in% seattletracts)

# Get population to calculate rates
schoolPop = c("B14001_007")
acsSchoolPop = get_acs(geography = "tract", variables = schoolPop, state = "53", county = "033", year = 2010, geometry = TRUE)
attach(acsSchoolPop)
dataSPop <- acsSchoolPop %>%
  filter(GEOID %in% seattletracts)
# Total enrollment
vecPop <- dataSPop$estimate 
# Above
vecPov1 <- dataSPov1$estimate
# Below
vecPov2 <- dataSPov2$estimate
# Rate Above
vecRate1 <- (vecPov1 / vecPop)
# Rate Below
vecRate2 <- (vecPov2 / vecPop)
dataSPov1$povRate <- vecRate1 * 100
dataSPov2$povRate2 <- vecRate2 * 100

# Make Maps
g1 <- dataSPov1 %>% 
  ggplot(aes(fill = povRate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Poverty Rates of High School Students", subtitle = "Above Poverty Line\n") +
  labs(fill = "Poverty Rate\n(2010 Census)")

g2 <- dataSPov2 %>% 
  ggplot(aes(fill = povRate2)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() +
  ggtitle("Poverty Rates of High School Students", subtitle = "Below Poverty Line\n") +
  labs(fill = "Poverty Rate\n(2010 Census)")

grid.arrange(g1, g2, nrow = 1)
```

Match Census tracts to districts:

```{r}
hsAttendanceZones <- read.csv("~/R/MATH 308 Data/High_School_Attendance_Areas_2022-2023.csv")
attach(hsAttendanceZones)
```

Educational Attainment Level:

```{r}
# Idea: Compare categorical with quantitative (educational attainment level vs. median income) Estimate change in median income based on educational attainment level
# Degree Level: 
# Less than a high school graduate
edcAttainment1 = c("S1501_C01_060E")
# High school graduate or equivalent (GED)
edcAttainment2 = c("S1501_C01_061E") 
# Some college or associates degree
edcAttainment3 = c("S1501_C01_062E") 
# Bachelors degree
edcAttainment4 = c("S1501_C01_063E") 
# Graduate or professional degree
edcAttainment5 = c("S1501_C01_064E") 

# Metropolitan statistical area
seattleCBSA = "42660"

#seattleEarnings = get_acs(geography = "cbsa", variables = edcAttainment1, year = 2010, geometry = TRUE)
#attach(seattleEarnings)
```






