---
title: "Shapefile Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(tigris)
library(tidyverse)
library(tidycensus)
library(ggplot2)
```

To-Do: Fix 15-19 student population

### Generate API Key for Census Data Scraping:

```{r, message=FALSE, warning = FALSE}
census_api_key("3488cfc51ef04c9886a285b2a7c770d4dcff78f8")
# To install API key for use in future sessions, run this function with `install = TRUE`
```

## Reading in Data

### Median Income by Tracts

```{r, message=FALSE, warning=FALSE}
medianIncomeVar = c("B19013_001")

# Median Household Income: 2006-2010 5-year ACS
acs1 = get_acs(geography = "tract", variables = medianIncomeVar, state = "53", county = "033", year = 2014, geometry = TRUE)
attach(acs1)
acs2 = get_acs(geography = "tract", variables = medianIncomeVar, state = "53", county = "033", year = 2019, geometry = TRUE)
attach(acs2)
pop1 = get_acs(geography = "tract", variables = "B14001_007E", state = "53", county = "033", year = 2014, geometry = TRUE)
attach(pop1)
pop2 = get_acs(geography = "tract", variables = "B14001_007E", state = "53", county = "033", year = 2019, geometry = TRUE)
attach(pop2)


# Read in list of Seattle block groups from SeattleGeoData
seattletracts <- read.csv("~/R/MATH_308_Data/Census_Tracts_2010.csv")  %>%
  pull("GEOID10") # Extracts a single column

# Subset data to Seattle Tracts
data10 <- acs1 %>%
  filter(GEOID %in% seattletracts) # Keep rows that match a condition
data14 <- acs2 %>%
  filter(GEOID %in% seattletracts)
data10p <- pop1 %>%
  filter(GEOID %in% seattletracts)
data10p <- data10p[-135,]
data14p <- pop2 %>%
  filter(GEOID %in% seattletracts)
data14p <- data14p[-135,]

master_data <- read_csv("~/Desktop/master_data.csv")
geo_guide <- data10 # Use census data for geography
geo_guide[8] <- master_data$house_price_med0914

to_check <- geo_guide$GEOID
cont_vect <- master_data$GEOID
missing <- c()
for (i in 1:135) {
  if (!(to_check[i] %in% cont_vect)) {
    append(missing, to_check[i])
  }
}
missing

```

### Maps and Visualization

```{r}
# Medium Household Income
p1 <- data10 %>%
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c(option = "magma") +
  ggtitle("King County By Median Household Income") +
  labs(fill = "Income Amount\n(2014 Census)")
p2 <- data14 %>%
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c(option = "magma") +
  ggtitle("King County By Median Household Income") +
  labs(fill = "Income Amount\n(2019 Census)")

# Tracts vs. School Districts
zones <- st_read("~/R/MATH_308_Data/High_School_Attendance_Areas_2022-2023.geojson")
attach(zones)
# Match coordinate systems
zones <- st_transform(zones, st_crs(data10))

p3 <- data10 %>%
  ggplot() +
  geom_sf(color = "grey50", fill = NA) + 
  geom_sf(data = zones, color = ("hotpink"), fill = "white", alpha = .3) +
  ggtitle("School District Area vs \n Census Tract")

# Median Value of Owner-Occupied Housing Units
p4 <- data14 %>%
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c(option = "magma") +
  ggtitle("King County By Median Household Income") +
  labs(fill = "Income Amount\n(2019 Census)")
```

## Manage Data

### Median Household Income by School District (2010-2014)

```{r}
# Match coordinate systems
crs1 <- st_crs(zones)
crs2 <- st_crs(data10)
zones <- st_transform(zones, st_crs(data10))


df_list_acs1 <- list()

# Specifically median income 
for (i in 1 : nrow(zones)) {
  data10 <- data10 %>%
  mutate(tractarea = st_area(.))
  
  thiszone <- zones %>%
    filter(HS_ZONE == HS_ZONE[i])
  df_name <- paste(HS_ZONE[i], "Data Frame")
  # Make the intersection
  intersectedzone <- thiszone %>%
    st_intersection(data10)
  # Split tract area at school zone boundary 
  zonearea <- intersectedzone %>%
  mutate(intersectedarea = st_area(.)) %>%
  mutate(weight = as.numeric(intersectedarea / tractarea)) %>%
  mutate(estimate_weighted = as.numeric(estimate * weight))
  
  assign(df_name, zonearea)
  df_list_acs1[[i]] <- zonearea
}

```

### Median Household Income by School District (2014-2019)

```{r}

# Match coordinate systems
crs1 <- st_crs(zones)
crs2 <- st_crs(data14)
zones <- st_transform(zones, st_crs(data14))


df_list_acs2 <- list()

# Specifically median income 
for (i in 1 : nrow(zones)) {
  data14 <- data14 %>%
  mutate(tractarea = st_area(.))
  
  thiszone <- zones %>%
    filter(HS_ZONE == HS_ZONE[i])
  df_name <- paste(HS_ZONE[i], "Data Frame (2)")
  # Make the intersection
  intersectedzone <- thiszone %>%
    st_intersection(data14)
  # Split tract area at school zone boundary 
  zonearea <- intersectedzone %>%
  mutate(intersectedarea = st_area(.)) %>%
  mutate(weight = as.numeric(intersectedarea / tractarea)) %>%
  mutate(estimate_weighted = as.numeric(estimate * weight))
  
  assign(df_name, zonearea)
  df_list_acs2[[i]] <- zonearea
}

```

### Population of School Age Children by School District (2010-2014)

```{r}

# Match coordinate systems
crs1 <- st_crs(zones)
crs2 <- st_crs(data10p)
zones <- st_transform(zones, st_crs(data10p))


df_list_pop1 <- list()

# Specifically median income 
for (i in 1 : nrow(zones)) {
  data10p <- data10p %>%
  mutate(tractarea = st_area(.))
  
  thiszone <- zones %>%
    filter(HS_ZONE == HS_ZONE[i])
  df_name <- paste(HS_ZONE[i], "Data Frame (3)")
  # Make the intersection
  intersectedzone <- thiszone %>%
    st_intersection(data10p)
  # Split tract area at school zone boundary 
  zonearea <- intersectedzone %>%
  mutate(intersectedarea = st_area(.)) %>%
  mutate(weight = as.numeric(intersectedarea / tractarea)) %>%
  mutate(estimate_weighted = as.numeric(estimate * weight))
  
  assign(df_name, zonearea)
  df_list_pop1[[i]] <- zonearea
}

```

### Population of School Age Children by School District (2014-2019)

```{r}

# Match coordinate systems
crs1 <- st_crs(zones)
crs2 <- st_crs(data14p)
zones <- st_transform(zones, st_crs(data14p))


df_list_pop2 <- list()

# Specifically median income 
for (i in 1 : nrow(zones)) {
  data14p <- data14p %>%
  mutate(tractarea = st_area(.))
  
  thiszone <- zones %>%
    filter(HS_ZONE == HS_ZONE[i])
  df_name <- paste(HS_ZONE[i], "Data Frame (4)")
  # Make the intersection
  intersectedzone <- thiszone %>%
    st_intersection(data14p)
  # Split tract area at school zone boundary 
  zonearea <- intersectedzone %>%
  mutate(intersectedarea = st_area(.)) %>%
  mutate(weight = as.numeric(intersectedarea / tractarea)) %>%
  mutate(estimate_weighted = as.numeric(estimate * weight))
  
  assign(df_name, zonearea)
  df_list_pop2[[i]] <- zonearea
}

```

### Median Value of Owner Occupied Houses

```{r}
housing["price_dif"] -> (housing2$estimate - housing$estimate)



# Match coordinate systems
crs1 <- st_crs(zones)
crs2 <- st_crs(data14p)
zones <- st_transform(zones, st_crs(data14p))


df_list_pop2 <- list()

# Specifically median income 
for (i in 1 : nrow(zones)) {
  data14p <- data14p %>%
  mutate(tractarea = st_area(.))
  
  thiszone <- zones %>%
    filter(HS_ZONE == HS_ZONE[i])
  df_name <- paste(HS_ZONE[i], "Data Frame (4)")
  # Make the intersection
  intersectedzone <- thiszone %>%
    st_intersection(data14p)
  # Split tract area at school zone boundary 
  zonearea <- intersectedzone %>%
  mutate(intersectedarea = st_area(.)) %>%
  mutate(weight = as.numeric(intersectedarea / tractarea)) %>%
  mutate(estimate_weighted = as.numeric(estimate * weight))
  
  assign(df_name, zonearea)
  df_list_pop2[[i]] <- zonearea
}

```

### Create Weighted Averages of Median Incomes by Population Density

```{r}
# Aggregate census statistics by school zone for all variables
# make a data fram with rows for each school district, add a column for weighted average of median income for each district
master_df <- data.frame(zones$HS_ZONE)
names(master_df)[1] <- "HS_ZONE"

med_av201014 <- c()

for (i in 1 : length(df_list_acs1)) {
  tot_value = sum(df_list_acs1[[i]]$estimate_weighted))
  med_av201014 = append(med_av201014, av_value)
}
master_df <- cbind(master_df, med_av201014)

med_av201419 <- c()

for (i in 1 : length(df_list_acs2)) {
  av_value = sum(df_list_acs2[[i]]$estimate_weighted) / nrow(df_list_acs2[[i]])
  med_av201419 = append(med_av201419, av_value)
}
master_df <- cbind(master_df, med_av201419)

pop_201014 <- c()

for (i in 1 : length(df_list_pop1)) {
  av_value = sum(df_list_pop1[[i]]$estimate_weighted) / nrow(df_list_pop1[[i]])
  pop_201014 = append(pop_201014, av_value)
}
master_df <- cbind(master_df, pop_201014)

pop_201419 <- c()

for (i in 1 : length(df_list_pop2)) {
  av_value = sum(df_list_pop2[[i]]$estimate_weighted) / nrow(df_list_pop2[[i]])
  pop_201419 = append(pop_201419, av_value)
}
master_df <- cbind(master_df, pop_201419)

```

# Median Household Value by Tract

```{r}
house_prices = c("B25077")

housing = get_acs(geography = "tract", variables = "B25077_001E", state = "53", county = "033", year = 2014, geometry = TRUE)

# Filter to Seattle
housing <- housing %>% 
  filter(GEOID %in% merge_attempt$GEOID)

to_merge <- data.frame(housing$GEOID, housing$estimate)
merge_attempt <- merge(merge_attempt, to_merge, by = "GEOID")

housing2 = get_acs(geography = "tract", variables = "B25077_001E", state = "53", county = "033", year = 2019, geometry = TRUE)

# Filter to Seattle
housing2 <- housing2 %>% 
  filter(GEOID %in% merge_attempt$GEOID)

to_merge2 <- data.frame(housing2$GEOID, housing2$estimate)
colnames(to_merge2)[1] = "GEOID"
colnames(to_merge2)[2] = "house_price_med1519"

merge_attempt <- merge(merge_attempt, to_merge2, by = "GEOID")

write.csv(merge_attempt, "/Users/Greta/Desktop/master_df.csv", row.names=TRUE)
```
